Predicting Ambient Air Pollution (PM2.5) Across the Contiguous U.S.

### 2023-04-23 Draft.Rmd
fghjk
Update Draft.Rmd

We used the random forest algorithm (RF) to determine which variables would be the best predictors for our model since RF is good at using variance to determine which features would be most influential for a model's predictive performance. In other words, the more variability there is in a dataset, the better RF performs at predicting which variables contribute to the outcome by averaging the results of all trees at the end and this averaging reduces the model's variance. Since our dataset has a wide variety of predictors ranging from education attainment levels to poverty levels in a given monitor region to emission data, RF would be good at reducing the variance in our dataset. For a given feature, the lower its impurity levels are, the more important that feature is. As a result, we got an ordered list of important variables and took the top 10 as input for the training set for our models. The top 10 important predictors were 'CMAQ', 'lat', 'county_area', 'lon', 'aod', 'popdens_county', 'zcta', 'log_nei_2008_pm25_sum_10000', and 'log_nei_2008_pm10_sum_10000'.

@@

### 3. Extreme Gradient Boost

>>>>

We used the random forest algorithm (RF) to determine which variables would be the best predictors for our model since RF is good at using variance to determine which features would be most influential for a model's predictive performance. In other words, the more variability there is in a dataset, the better RF performs at predicting which variables contribute to the outcome by averaging the results of all trees at the end and this averaging reduces the model's variance. Since our dataset has a wide variety of predictors ranging from education attainment levels to poverty levels in a given monitor region to emission data, RF would be good at reducing the variance in our dataset. For a given feature, the lower its impurity levels are, the more important that feature is. As a result, we got an ordered list of important variables and took the top 10 as input for the training set for our models. The top 10 important predictors were 'CMAQ', 'lat', 'county_area', 'lon', 'aod', 'popdens_county', 'zcta', 'log_nei_2008_pm25_sum_10000', and 'log_nei_2008_pm10_sum_10000'. This website was used [TowardDataScience](https://towardsdatascience.com/feature-selection-using-random-forest-26d7b747597f).

@@

### 3. Extreme Gradient Boost Regression
The Extreme Gradient (XG) Boosting model performs by training weak learners, models that have low prediction accuracy, sequentially to become a single strong learner, a model that has strong prediction accuracy. We preprocess the training set to make the predictive modeling process more accurate and smoother by prepping (estimates the quantities for 'train_scaled') and baking (assigns prepped data to an object) the data. The aforementioned top 10 predictors were used in this preprocessing recipe. Moreover, the categorical variables were factored and features with no variance were removed before this data was baked. Then, this processed data is assigned to 'folds3' and resampled on 5 subsets of the training set. We tune the hyperparameters to make the modeling process more efficient and perform grid specification to determine which hyperparameter values have high prediction accuracy (i.e. low prediction error). Next, we isolated the best hyperparameter values, so they can be used for the final boosting model. Lastly, we fitted the model and evaluated its prediction performance. Its RMSE score was .71, indicating the XG Boosting is a good model.
